A Systematic Approach to Analyzing Syndication of Viral News Stories

### 2019-12-07 paper/paper.tex
Update on Overleaf.

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}
\usepackage{array}
\usepackage{longtable}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{wrapfig}
\usepackage[dvipsnames]{xcolor}
\title{A Systematic Approach to Analyzing Syndication of Viral News Stories}

@@

\begin{abstract}
In this work, we present a novel, semiautonomous methodology for analyzing syndicated news articles. We first detail our manual process of news syndication analysis from which we observe a number of unsettling patterns. From there, we go on to categorize and define two different forms of syndication: top-down and bottom-up. We then transform our manual process into a semiautomated one which combines web scraping and machine learning techniques to create and then analyze a data set of viral news stories. From the analysis, we are able to make some inferences about the nature of syndicated news articles, as well as trace back a viral news story to its original publisher and article. By presenting this work, we hope to minimize the effects that blind news syndication has on the spread of misinformation. 
\end{abstract}
\keywords{Fake news \and News propagation \and News syndication}
\newpage
\tableofcontents
\newpage
\section{Introduction}
Misinformation, disinformation, and fake news have become a common occurrence in our interconnected society. The idea of intentionally spreading fake news to sway public opinion, also known as disinformation, was brought to light by the U.S. Intelligence Committee's report on Russian meddling in the 2016 U.S. presidential election \cite{stelzenmuller2017impact}. Misinformation, or the unintentional spread of inaccurate news, has also become ubiquitous thanks to social media, which allows unreliable news to be proliferated quickly and without any verification \cite{pourghomi2017stop}. The use of a combination of misinformation and disinformation by nation states, non-state actors, and other politically motivated groups has become so prevalent that it has affected over seventy countries \cite{alba_satariano_2019}. This problem has gotten so extreme that the integrity of the U.S. election system, and even the fundamental democratic ideals of the U.S., have been called into question by some \cite{mueller2019report}. Due to the global nature of this problem, over fifty countries have begun to implement measures to combat misinformation and disinformation \cite{poynter}. The measures implemented by these countries span from laws and legislation, to creating task forces and launching official investigations to try to weed disinformation \cite{poynter}. Besides nation states trying to tackle this behemoth of a problem, many researchers and academics have also been trying to implement various methods to try to detect and classify misinformation, disinformation, and fake news. This area of research has become commonly known as \textit{fake news detection} \cite{zhang2019overview}.

\subsection{Problem Statement}
There has already been many different approaches proposed to detecting and classifying fake news. Many proposed solutions have mostly relied on various machine learning techniques, such as using deep neural networks, or recurrent learning models \cite{shu2017fake}. These models for the most part work by training the model on a variety of features ranging anywhere from text-based features, to network features that show how the article has been spread through space and time \cite{shu2017fake}. The problem with these models is that they rely on a lot of information to be fed into them, which take a long time to extracted. Also, many of these approaches have only been tested on a few media outlets and social media platforms and could not work on a global scale.

@@

\subsection{Contributions}
We make a number of contributions throughout this work. First, we present out initial manual methodology which outlines how we first identified viral articles, how we traced them back to their origin, and the findings from out analysis. From this analysis, we present and define two different types of news syndication: top-down and bottom-up. We then present a refined, semiautonomous methodology which is derived from our initial methodology, but with a few revisions. This methodology, given a viral news story and a couple related articles, will automatically scrape the web and identify articles within a set time window which were either syndicated from the original set of articles fed to it, or articles that the original set of articles syndicated. We use a combination and similarity score and key words that eventually feed into a dendrogram which clusters similar articles together. We then take this clustered data set and build a timeline which allows the original article to be identified. We also are able to perform a wide range of analysis on various article features and categorize how certain features change over time. This analysis then allows us to identify patterns about syndicated articles and eventually draw conclusions.

@@

The main difference between these definitions is the intent and what it is being used for. These terms may be used interchangeably in this work. This is because our work does not intend to differentiate between the intent of the news. Our work focuses on building a methodology that is able to semiautonomously trace back an article to its origin and analyze how article features change over time. Future work involves trying to classify news as fake or not, but that involves using already developed models for classification. Not even these models attempt to differentiate between different intentions. Therefore, although misinformation, disinformation, and fake news do technically have different meanings, they can be used interchangeably in this line of work since most research on fake news detection just attempts to classify news as accurate or inaccurate, not to try to determine the intent behind the article.

@@

%But what exactly is fake news? According to Wikipedia, "fake news is a form of news consisting of deliberate disinformation or hoaxes spread via traditional news media (print and broadcast) or online social media". In addition to this definition, fake news may also include clickbait stories, propaganda, satire, biased news, or even sometimes just bad journalism.

@@

\subsection{Case Study: University of Virginia}
Once we found the story about the Chinese gold story, we wanted to see if other similar articles exhibited the same properties and characteristics. We started looking for articles which origin could be traced back to either a social media post, or a local news media outlet. One of the first articles that we identified was the story titled \texit{University of Virginia cancels 21-gun salute from Veterans Day ceremony} on Fox News.

\subsection{Findings}
We found that

@@

@@

The approach can be broadly categorized into two stages.
\begin{enumerate}
  \item 
  \textit{Dataset Generation}: As a part of this data generation we identify the  viral news articles, scrape for articles that are similar to it and gather metadata on these set articles. The raw data is passed through machine learning models to group them as relevant articles and discard non relevant articles.
   \item
   \textit{Analysis}: Numerous data visualization methods were applied based on the findings of phase I. These included time series analysis and histograms on certain features.
\end{enumerate}

\subsection{Dataset Creation}
\subsubsection{Identification of Viral News Article}
Like in phase I, we did not use articles published by unbiased media sources and steered clear of stories related to Donald Trump. An article with more than \emph{ten thousand reactions}\footnote{reaction is the sum of likes, comments and shares} on facebook was quantified as a viral news article. The articles selected in the end were a heterogenous mixture of local and international events. These new articles were used to manually build\emph{url.csv} dataset. This dataset consists of a single feature, url pointing to the article.

\subsubsection{Scraping similar news articles}
Once a url.csv is generated, remanining articles are fetched manually of the internet. Currently, we use google news api and newsapi to query for the relevant articles.
The urls in \emph{url.csv} were used to generate a temporary dataset which consists of the url of the story along with the title of the story. The title is

\subsubsection{Building Dataset}
\colorbox{BurntOrange}{Add code snippet here}

\subsection{Data Clustering}
\subsubsection{Similarity Score}
\subparagraph{Stemming}
\subparagraph{TF-IDF}
\subparagraph{Cosine Distance}
\subsubsection{Hierarchical Clustering}
\subsection{Data Analysis}
\subsection{Case Study : The Arkansas Story}
\subsubsection{Overview}
\subsubsection{Findings}

>>>>

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}
\usepackage{array}
\usepackage{longtable}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{wrapfig}
\title{A Systematic Approach to Analyzing Syndication of Viral News Stories}

@@

\begin{abstract}
In this work, we present a novel, semiautonomous methodology for analyzing syndicated news articles. We first detail our manual process of news syndication analysis from which we observe a number of unsettling patterns. From there, we go on to categorize and define two different forms of syndication: top-down and bottom-up. We then transform our manual process into a semiautomated one which combines web scraping and machine learning techniques to create and then analyze a data set of viral news stories. From the analysis, we are able to make some inferences about the nature of syndicated news articles, as well as trace back a viral news story to its original publisher and article. By presenting this work, we hope to minimize the effects that blind news syndication has on the spread of misinformation. 
\end{abstract}
\keywords{Fake news \and News propagation \and News syndication}
\newpage
\tableofcontents
\newpage
\section{Introduction}
Misinformation, disinformation, and fake news have become a common occurrence in our interconnected society. The idea of intentionally spreading fake news to sway public opinion, also known as disinformation, was brought to light by the U.S. Intelligence Committee's report on Russian meddling in the 2016 U.S. presidential election \cite{stelzenmuller2017impact}. Misinformation, or the unintentional spread of inaccurate news, has also become ubiquitous thanks to social media, which allows unreliable news to be proliferated quickly and without any verification \cite{pourghomi2017stop}. The use of a combination of misinformation and disinformation by nation states, non-state actors, and other politically motivated groups has become so prevalent that it has affected over seventy countries \cite{alba_satariano_2019}. This problem has gotten so extreme that the integrity of the U.S. election system, and even the fundamental democratic ideals of the U.S., have been called into question by some \cite{mueller2019report}.

But what exactly is fake news? According to Cambridge Dictionary \cite{FAKENEWS51:online}, fake news is "false stories that appear to be news, spread on the internet or using other media, usually created to influence political views or as a joke". In addition to this definition, fake news may also include clickbait stories, propaganda, satire, biased news, or even sometimes just bad journalism \cite{Explaine83:online}.

These days, anyone can publish content on the internet, whether it is on a website, a blog, or a social media profile, and potentially reach large audiences. The fact that people nowadays rely on the internet to consume news doesn't make this problem any easier to solve.
Fake news is very profitable for the publisher \cite{Explaine83:online}, as viral stories generate large amounts of traffic, thus increasing advertisement revenue. For that reason, fake news are widely common around the internet. 
Another motive to publish fake news is to promote a political agenda. Fake political news stories are meant to persuade consumers to accept a biased of false beliefs. This is the main reason detecting fake news is such a challenge. These false stories were created to sound true and interesting.

Due to the global nature of this problem, over fifty countries have begun to implement measures to combat misinformation and disinformation \cite{poynter}. The measures implemented by these countries span from laws and legislation, to creating task forces and launching official investigations to try to weed disinformation \cite{poynter}. Besides nation states trying to tackle this behemoth of a problem, many researchers and academics have also been trying to implement various methods to try to detect and classify misinformation, disinformation, and fake news. This area of research has become commonly known as \textit{fake news detection} \cite{zhang2019overview}.

\subsection{Problem Statement}
One can try and detect fake news by checking the source of the story, read beyond the title (which is intended to attract users and thus contains more interesting and controversial topics), try to trace the article back to the original source, or even make sure the article is not a satire by any chance \cite{Explaine83:online}.
Even though it's possible, doing so is not an easy task, and that is why there has already been many different approaches proposed to detecting and classifying fake news. Many proposed solutions have mostly relied on various machine learning techniques, such as using deep neural networks, or recurrent learning models \cite{shu2017fake}. These models for the most part work by training the model on a variety of features ranging anywhere from text-based features, to network features that show how the article has been spread through space and time \cite{shu2017fake}. The problem with these models is that they rely on a lot of information to be fed into them, which take a long time to extracted. Also, many of these approaches have only been tested on a few media outlets and social media platforms and could not work on a global scale.

@@

\subsection{Contributions}
We make a number of contributions throughout this work. First, we present out initial manual methodology which outlines how we first identified viral articles, how we traced them back to their origin, and the findings from out analysis. From this analysis, we present and define two different types of news syndication: top-down and bottom-up. We then present a refined, semi-autonomous methodology which is derived from our initial methodology, but with a few revisions. This methodology, given a viral news story and a couple related articles, will automatically scrape the web and identify articles within a set time window which were either syndicated from the original set of articles fed to it, or articles that the original set of articles syndicated. We use a combination and similarity score and key words that eventually feed into a dendrogram which clusters similar articles together. We then take this clustered data set and build a timeline which allows the original article to be identified. We also are able to perform a wide range of analysis on various article features and categorize how certain features change over time. This analysis then allows us to identify patterns about syndicated articles and eventually draw conclusions.

@@

The main difference between these definitions is the intent and what it is being used for. These terms may be used interchangeably in this work. This is because our work does not intend to differentiate between the intent of the news. Our work focuses on building a methodology that is able to semi-autonomously trace back an article to its origin and analyze how article features change over time. Future work involves trying to classify news as fake or not, but that involves using already developed models for classification. Not even these models attempt to differentiate between different intentions. Therefore, although misinformation, disinformation, and fake news do technically have different meanings, they can be used interchangeably in this line of work since most research on fake news detection just attempts to classify news as accurate or inaccurate, not to try to determine the intent behind the article.

@@

%But what exactly is fake news? According to Wikipedia \cite{Fakenews55:online}, "fake news is a form of news consisting of deliberate disinformation or hoaxes spread via traditional news media (print and broadcast) or online social media". In addition to this definition, fake news may also include clickbait stories, propaganda, satire, biased news, or even sometimes just bad journalism \cite{Explaine83:online}.

@@

\subsection{Case Study: University of Virginia}
Once we found the story about the Chinese gold story, we wanted to see if other similar articles exhibited the same properties and characteristics. We started looking for articles which origin could be traced back to either a social media post, or a local news media outlet. One of the first articles that we identified was the story titled \textit{University of Virginia cancels 21-gun salute from Veterans Day ceremony} from Fox News.

\begin{figure}[h]
    \centering
    \includegraphics[width=16.5cm]{paper/img/uva-story.png}
    \caption {First Article About UVA That We Found}
    \label{fig:uvastory}
\end{figure}

\subsubsection{Propagation Analysis}
When we traced the propagation path back to the original published article, it turned out that it was actually an opinion piece from a local newspaper which is located around the University of Virginia. Once again, we had to use a manual process to trace back the article to its origin since we started with the Fox News article, which had already been syndicated multiple times.

\begin{figure}[h]
    \centering
    \includegraphics[width=16.5cm]{paper/img/uva-timeline.png}
    \caption {Timeline of UVA Story}
    \label{fig:uvatimeline}
\end{figure}
Analyzing the propagation path, it took about five days for the story to move from the local newspaper, to national news. It is difficult to compare this timeline to that of the Chinese Gold story since that story was international in scope and this story is local to the United States. Nevertheless, it is interesting to note the time delay between when the article was first posted by the local newspaper, and when a national media outlet, such as Fox News, syndicated the article.

\subsubsection{Article Feature Analysis}
We wanted to see if the general trends from the Chinese gold story of article length increasing over time and title length increasing over time would hold true with this article. If they did, this would be an interesting correlation that could be drawn between syndicated articles that originate from local news outlets, or social media platforms.

\begin{figure}[h]
\centering
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width=7cm]{paper/img/uva-articlelen.png}
  \captionof{figure}{Article Length Over Time}
  \label{uvaarticlelen}
\end{minipage}%
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width=7cm]{paper/img/uva-titlelen.png}
  \captionof{figure}{Title Length Over Time}
  \label{uvatitlelen}
\end{minipage}
\end{figure}

Unlike the Chinese gold story, there was no clear correlation between time since original publication of the article, and the article length. However, the length of every syndicated article was longer than the original. This same pattern can be seen with the title length over time; every syndicated article title length was longer than the original article's title length.

\subsection{Findings}
The findings from our manual analysis are

@@

\subsection{Limitations of a Manual Approach}

@@
